---
title: "cluster_iris"
author: "zfwang"
date: "2019年1月14日"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# **数据**

```{r}
# data
iris = datasets::iris

# split data and label
iris2 = iris[, -5]
species_labels = iris[, 5]
head(iris2)
species_labels
```


# **自变量散点图矩阵**

```{r}
# get nice colors
library(colorspace)
species_col = rev(rainbow_hcl(3))[as.numeric(species_labels)]

# Plot a SPLOM
pairs(iris2, 
      col = species_col,
      lower.panel = NULL,
      cex.labels = 2,
      pch = 19,
      cex = 1.2)

# Add a legend
par(xpd = TRUE)
legend(x = 0.05,
       y = 0.4,
       cex = 1,
       legend = as.character(levels(species_labels)),
       fill = unique(species_col))
par(xpd = NA)
```

> * 从散点图矩阵中可以看出，`Setosa`和`Versicolor`、`Virginica`明显不同：`Setosa`有较小的花瓣(petal)长度和宽度，而`Versicolor`、`Virginica`两种花不容易根据他们的花萼和花瓣的长度和宽度分离开来


# **花萼、花瓣的平行坐标图(parallel coordinates plot)**


```{r}
par(las = 1, 
    mar = c(4.5, 3, 3, 2) + 0.1, 
    cex = 0.8)
MASS::parcoord(iris2, 
               col = species_col, 
               var.label = TRUE, 
               lwd = 2)

## Add Title
title("Parallel coordinates plot of the Iris data")

# Add a legend
par(xpd = TRUE)
legend(x =1.75, 
       y = -0.15, 
       cex = 1,
       legend = as.character(levels(species_labels)),
       fill = unique(species_col),
       horiz = TRUE)
```


> 从上面的平行坐标图中也可以看出，`Setosa`有较小的花瓣(petal)长度和宽度，而`Versicolor`、`Virginica`两种花不容易根据他们的花萼和花瓣的长度和宽度分离开来


# **层次聚类(Hierarchical Clustering)**

* hierarchical clustering method: "complete"
* cluserting data visualize

```{r}
# 计算距离
d_iris = dist(iris2)

# 层次聚类
hc_iris = hclust(d_iris, method = "complete")
iris_species = rev(levels(species_labels))
```


```{r}
library(dendextend)

dend = as.dendrogram(hc_iris) %>%
    rotate(1:150) %>%
    color_branches(k = 3)

labels_colors(dend) = rainbow_hcl(3)[sort_levels_values(as.numeric(iris[, 5][order.dendrogram(dend)]))]
labels(dend) = paste(as.character(iris[, 5])[order.dendrogram(dend)], "(", labels(dend), ")", sep = "")
dend = hang.dendrogram(dend, hang_height = 0.1)
dend = set(dend, "labels_cex", 0.5)

# Plot
par(mar = c(3, 3, 3, 7))
plot(dend, 
     main = "Clustered Iris data set(the labels give the true flower species)",
     horiz = TRUE, 
     nodePar = list(cex = 0.007))
legend("topleft", legend = iris_species, fill = rainbow_hcl(3))
```


```{r}
par(mar = rep(0, 4))
circlize_dendrogram(dend)
```












